### 2015-08-27 rand_forest.rmd
tried boosted trees

---
title: "Regression"
author: "Amy Cook"
date: "August 15, 2015"
output: word_document
---

```{r, echo=FALSE, include=FALSE}

library('knitr', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.2/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.2/library')
library('caret',lib='C:/Progra~1/R/R-3.2.2/library')
library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
library('rpart',lib='C:/Progra~1/R/R-3.2.2/library')
library('rpart.plot',lib='C:/Progra~1/R/R-3.2.2/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.2/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.2/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.2/library')

library(devtools)
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')
library('randomForest')
library('caret')

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d$Post.Code<- as.factor(all7d$Post.Code)

```

This file is to do a multi-step regression to predict RETURN PER DOLLAR.
The reason for multistepping is to allow for multiple variables with lots of missing NA values.

Variables to include:

Full:

inv.mlsto
timespan
no.users
Discipline
pc.pro
client.totinv
code.director
Num.disc
Business
Biz.type
code.ProjEng

Many NA:

code.contact
code.client
Project.Value
JD.Second
Post.Code
Billing.Type
majority.pos
pc.majpos

The 'Many NA' variables will be split into groups:

* code.contact: 1652 NA
* code.client: 422 NA
* Project.Value: 2333 NA
* JD.Second: 1057 NA
* Post.Code: 356 NA
* Billing.Type: 1314 NA
* majority.pos: 377 NA
* pc.majpos: 377 NA

Will address each of the variables with many NA separately. 8 of these variables totally. run a separate lm for each of these

No need to transform numeric variables - not required for trees

Root mean square error as initial check.

```{r, echo=FALSE}

#predict Petal.Width using lm, find RMSE
#answer will be return.pdol
#guess will be response of prediction which is cube root
RMSE<- function(answer, guess){
        error= answer-guess
        root.error= sqrt(mean(error^2))
        print(root.error)
}

```

Now fit full variables to random forest, tune mtry and number of trees

Try rpart first

```{r, echo=FALSE}

train_ind <- sample(seq_len(nrow(all7d)), size = nrow(all7d)*2/3)

train <- all7d[train_ind, ]
test <- all7d[-train_ind, ]

#create tree with training data only
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
# summary(tree.all7d)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d)
plotcp(tree.all7d)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train,
                  control = rpart.control(cp= 0.013))

split.fun <- function(x, labs, digits, varlen, faclen){
  # replace commas with spaces (needed for strwrap)
  labs <- gsub(",", " ", labs)
  for(i in 1:length(labs)) {
    # split labs[i] into multiple lines
    labs[i] <- paste(strwrap(labs[i], width=25), collapse="\n")
    }
  labs
  }

prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 49.8 cents, already close to regression..

#try to make a rainbow tree
heat.tree <- function(tree, low.is.green=FALSE, ...) { # dots args passed to prp
  y <- tree$frame$yval
  if(low.is.green)
    y <- -y
  max <- max(y)
  min <- min(y)
  cols <- rainbow(99, end=.36)[
    ifelse(y > y[1], (y-y[1]) * (99-50) / (max-y[1]) + 50,
           (y-min) * (50-1) / (y[1]-min) + 1)]
  prp(tree, branch.col=cols, box.col=cols, ...)
  }

heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)

```

RMSE is 0.498 which is worse than regression in predicting return per dollar.

Decision trees not affected by too many variables - leave all in

Decision trees cannot handle pairwise interactions

Let's try 10 fold cross validation and have a look at the error.
Use caret package

```{r, echo=FALSE}
#10 fold cross validation using caret package
#specify 10 fold
tc <- trainControl("cv",10)
#keep cp static at 0.011
rpart.grid <- expand.grid(.cp=0.011)
train.rpart <- train( return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=all7d, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

#RMSE = 0.505
# Rsquared = 0.095, quite bad

```

Let's try adding more variables:

```{r, echo=FALSE}
#try adding code.client

addone<- function(var= 'JD.Second', min.Year = 2008, df= all7d){
        reduced = df[!(is.na(df[,var])) & df$Year > min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type", var, sep=" + ")
        add.one = rpart(as.formula(formula), data = reduced)

        #select cp
        cp.new = add.one$cptable[which.min(add.one$cptable[,"xerror"]),"CP"]-0.0002
        add.one = rpart(as.formula(formula), data = reduced,
                        control = rpart.control(cp= cp.new))

        #plot decision tree
        print(
          heat.tree(add.one, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
          )

        #perform 10 fold xval using caret package
        tc = trainControl("cv",10)
        #keep cp static at 0.011
        rpart.grid = expand.grid(.cp=cp.new)
        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

        print(
        train.rpart$results
        )
}

#without additional variables, RMSE sits at 0.4759

addone(var='JD.Second', min.Year = 2008, df=all7d)
#RMSE = .528
addone(var='Post.Code', min.Year = 2008, df=all7d)
#RMSE = .499
addone(var='majority.pos', min.Year = 2008, df=all7d)
#RMSE = .49
addone(var='pc.majpos', min.Year = 2008, df=all7d)
#RMSE = .486
addone(var='code.client', min.Year = 2008, df=all7d)
#RMSE = .534
addone(var='Billing.Type', min.Year = 2008, df=all7d)
#RMSE = .502
addone(var='code.contact', min.Year = 2008, df=all7d)
#RMSE = .535

# none of the extra variables actually appear in the revised trees except JD.Second
# can only assume the different subsets are what is causing the variation in RMSE

```

Write a function that can plot a tree
Limit Year
choose additional variables to include.

```{r, echo=FALSE}

addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type")

        for(i in 1:length(var)){
                formula = paste(formula, var[i], sep= " + ")
        }

        add.mult = rpart(as.formula(formula), data = reduced)

        #select cp
        cp.new = add.mult$cptable[which.min(add.mult$cptable[,"xerror"]),"CP"]-0.0002
        add.mult = rpart(as.formula(formula), data = reduced,
                        control = rpart.control(cp= cp.new))

        #plot decision tree
        print(
          heat.tree(add.mult, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
          )

        #perform 10 fold xval using caret package
        tc = trainControl("cv",10)
        #keep cp static at 0.011
        rpart.grid = expand.grid(.cp=cp.new)
        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

        print(
        train.rpart$results
        )

}

#try Billing.Type with inv.mlsto.log
## YES! print this one!! 
addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
#RMSE at 48.6 cents

```

Adding aditional variables is making everything worse.
Try randomForest

```{r, echo=FALSE}
#perform 10 fold xval using caret package
tc = trainControl("cv",5)
rf.grid = expand.grid(mtry= 1:10)
min.form= "return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type"
train.rf = train(as.formula(min.form), data=all7d, method="rf",trControl=tc,
                    tuneGrid=rf.grid)
#result - use mtry= 5

#make first single randomForest
first.rf<- randomForest(as.formula(min.form),
                  data=all7d,mtry=1,ntree=500,importance=TRUE)

#already have made train and test set
#try to predict medv
rf.all<-randomForest(as.formula(min.form),data=all7d,mtry=1,ntree=500,importance=TRUE)

#create vectors full of '0's
oob.err<-double(7)
test.err<-double(7)
for(mtry in 1:7){
        fit<-randomForest(as.formula(min.form),data=train, mtry=ntry, ntree=500)
        oob.err[mtry]<-fit$mse[mtry]
        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
        cat(mtry," ")
}

qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(1:mtry,oob.err),colour="blue")+
        geom_point(aes(1:mtry,oob.err),colour="blue")

#mtry = 3 is optimal! test RMSE of 48.2 cents ....
oob.err<-double(5)
test.err<-double(5)
ntree = c(300,400,500,600,700)
for(i in 1:5){
        fit<-randomForest(as.formula(min.form),data=train, mtry=3, ntree=ntree[i])
        oob.err[i]<-fit$mse[i]
        test.err[i]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
        cat(i," ")
}

qplot(ntree,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(ntree,oob.err),colour="blue")+
        geom_point(aes(ntree,oob.err),colour="blue")
#doesn't really matter, use 400!

#final forest with full variables

full.forest<- randomForest(as.formula(min.form),data=all7d,mtry=5,ntree=500,importance=TRUE)
plot(full.forest)
importance(full.forest)
varImpPlot(full.forest)

full.forest<- randomForest(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type +code.client, data= all7d[complete.cases(all7d$code.client),],
                  mtry=5,ntree=500,importance=TRUE)

```

Try adding variables

```{r, echo=FALSE}

addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type")

        for(i in 1:length(var)){
                formula = paste(formula, var[i], sep= " + ")
        }

        add.mult = randomForest(as.formula(formula),mtry=5, ntree=500, data = reduced)

        #perform 10 fold xval using caret package
        tc = trainControl("cv",5)
        #keep cp static at 0.011
        rf.grid = expand.grid(mtry= 5:6)
        train.rf = train(as.formula(formula), data=df, method="rf",trControl=tc,
                  tuneGrid=rf.grid)

        print(
        train.rf
        )

}

#try Billing.Type with inv.mlsto.log
## YES! print this one!! 
addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
#RMSE at 48.6 cents

```

Trial some scenarios:

Run full variables plus

Make a function to output:

* var importances
* RMSE

based on a list of variables that you want to use

```{r, echo=FALSE}

sample<- all7d[1000,]
sample<- sample[,!is.na(sample)]
vars<- names(sample) 
vars<- vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc','Biz.type','Post.Code',
               'code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto','balance.mlsto','client.totinv.log','pc.majpos')]
vars

RMSE<- function(answer, guess){
        error= answer-guess
        root.error= sqrt(mean(error^2))
        # print(root.error)
}

final.output<- function(df=all7d, samp = 800) {
        #create sample df
        sample = df[samp,]
        sample = sample[,!is.na(sample)]
        vars= names(sample)
        vars = vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc',
                                  'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
                                  'balance.mlsto','client.totinv.log','pc.majpos')]
        variables = vars

        print(
                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log, -pc.majpos.log)))
        )

        #make reduced dataset with complete cases for all variables we want to use
        reduced = df[complete.cases(df[,variables]),]

        print(
                paste(nrow(reduced), "cases in full model")
        )

        formula = paste("return.pdol ~ ")

        for(i in 1:length(variables)){
                formula = paste(formula, variables[i], sep= " + ")
        }

        if('inv.mlsto.log' %in% variables &
           'inv.mlsto.log' %in% variables){
                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
        }

        if('inv.mlsto.log' %in% variables &
           'code.client' %in% variables){
                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
        }

        if('JD.Second' %in% variables &
           'no.users' %in% variables){
                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
        }

        final = aov(as.formula(formula), data = reduced)

        fit= summary(final)

        ans= data.frame(answer= reduced$return.pdol, final.lm = predict(final, reduced, interval= 'predict'))

        # print(head(ans))

        #print variable importances for reduced lm just run

        imp= data.frame('vars'= rownames(fit[[1]]), 'F.val' = fit[[1]]$`F value`)
        #remove white space from levels in imp$vars
        levels(imp$vars)<- sub("\\s+$", "", levels(imp$vars))
        #remove Residuals row
        imp= imp[!(imp$vars %in% 'Residuals'),]

        #print variable importances
        print(
                list("variable importances",
                     imp %>% arrange(-F.val) %>% head())
        )

        #print RMSE for cases that match certain categories only
        #first subset predictions to cases that match the category

        reduced$predict= ans$final.lm.fit

        cats= c('code.client', 'code.contact', 'JD.Second', 'Business', 'Discipline', 'majority.pos')

        for(i in 1:length(cats)){
                if(cats[i] %in% vars){
                        cats.red = reduced[reduced[,cats[i]] %in% sample[,cats[i]],]
                        cat.RMSE = RMSE(cats.red$return.pdol, cats.red$predict) %>% round(4)
                        print(
                                paste(cats[i], "RMSE =", cat.RMSE, sep= " "))
                        print(
                                paste("Number of cases =", nrow(cats.red), sep= " "))
                        print(
                                head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto','return.pdol','predict',cats[i])])
                        )
                        }
        }

        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(4)

        print(
        paste("Overall predicted RMSE=", final.RMSE , sep= " ")
        )

        predict.rpdol = predict(final, sample) %>% round(3)
        predict.rpdol = predict.rpdol[[1]]
        predict.cost = sample$inv.mlsto/(1+predict.rpdol) %>% round(1)
        print(
               paste("return per dollar prediction", predict.rpdol, sep = " ")
        )

        print(
               paste("answer", sample$return.pdol %>% round(3), sep = " ")
        )

        print(
                paste("Invoiced= $", sample$inv.mlsto, ",   predicted cost= $", predict.cost %>% round(1), sep="")
        )

        print(
                paste("Actual cost= $", sample$cost.mlsto, sep="")
        )

}

final.output(df=all7d, samp = 500)

```