### 2015-10-30 financial.rmd
finish forest and log

---
title: "Financial_Analysis"
author: "Amy Cook"
date: "October 21, 2015"
output: word_document
---

#Assess financial implications for a predictive model

We typically get a confusion matrix built from probabilities

```{r_loadpackages, echo=FALSE, include=FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')

# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

Load data

```{r, echo=FALSE}
test.pred <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
all7d <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')
test.pred<- merge(test.pred, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")

```

Create vector of profit points

```{r, echo=FALSE}

threshold<- seq(0,1, by=.05)
fin.prof<- rep(0, length(threshold))

for(i in 1:length(threshold)){
        prof<- data.frame("balance" = test.pred$balance.mlsto, 
                  "pred" = test.pred$pred,
                  "ans" = test.pred$b.rpdol)
        #pred = probabilities
        prof$conf <- rep(0, nrow(prof))
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
        prof$profit<- rep(0, nrow(prof))
        prof$profit<- ifelse(prof$conf %in% "TP", -prof$balance, prof$profit)
        prof$profit<- ifelse(prof$conf %in% "FP", -prof$balance, prof$profit)

        fin.prof[i] = sum(prof$profit)

}

prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)

ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()

# if say no to all jobs with higher than prob of 0.5, save $500,000

head(prof)

####
# SUMMARY STATS
####

# would dismiss 56 jobs
prof %>% filter(pred>=0.5) %>% nrow

# would dismiss 19 profitable jobs
prof %>% filter(pred >= 0.5 & balance > 0) %>% nrow
#out of 607
prof %>% filter(balance > 0) %>% nrow

# would dismiss 37 losing jobs
prof %>% filter(pred >= 0.5 & balance < 0) %>% nrow
#out of 135
prof %>% filter(balance < 0) %>% nrow

#without algorithm, balance would be $1,475,642
sum(prof$balance)

# without any losing jobs, balance would be $2,516,246
prof %>% filter(balance > 0) %>% sum

# without any profitable jobs, balance would be -$1,040,322
prof %>% filter(balance < 0) %>% sum

#how much real profit would you be dismissing? $82631
prof.dismis<- prof %>% filter(pred >= 0.5 & balance > 0) %>% select(balance) %>% sum
#how much real loss would you be avoiding? $566,481.6
loss.dismis<- prof %>% filter(pred >= 0.5 & balance < 0) %>% select(balance) %>% sum
# total benefit = $483,850
algor.prof<- -(loss.dismis + prof.dismis)

# total balance would be $1,959,492
sum(prof$balance) + algor.prof

# a 32.8% improvement
algor.prof/sum(prof$balance)

```