### 2016-05-26 model_averaging.rmd
backwards at work

---
title: "Model Averaging"
author: "Amy Cook"
date: "May 26, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('caret')
library('pROC')
library('e1071')
library('randomForest')
library('party')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]

```

Model averaging how to do it>?
Best models are ..

* boosted tree
* logisitic regression
* random forest

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.