### 2016-05-27 rf_log_naiive_auc.rmd
re-syncing mac draft

---
title: "test_imputed"
author: "Amy Cook"
date: "October 27, 2015"
output: word_document
---

Test on random forest, logistic regression and naiive bayes :)
Compare:

* subset 9c data
* rf imputed
* rfMICEd imputed
* subset 9a data (uncompressed JD.Second and Business)
* rf imputed with 9a
* rfMICE imputed with 9a

Out of the best from above, can experiment with adding in client and contact.

Load 3 datasets:

```{r, echo=FALSE}
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('caret')
library('pROC')
library('e1071')
library('randomForest')
library('party')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]
all10rf<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10rf.csv')[,-1]
all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
all9c<- read.csv('all9c.csv')[,-1]
all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]

#VARS::::
# no.users, Discipline, pc.pro, Business, code.client,JD.Second, majority.pos, b.timespan.cbrt,
# inv.mlsto.log, client.totinv.log, pc.majpos.log, code.contact, Billing.Type, mlsto, Year, b.rpdol

all10a<- all9c %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                         mlsto, Year, b.rpdol)
all10a<- all10a[complete.cases(all10a),]
# 2280 cases - 10 vars

all10b<- all9c %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10b<- all10b[complete.cases(all10b),]
# 1987 cases - 12 vars

all10c<- all9c %>% select(no.users, Discipline, pc.pro, Business, JD.Second,
                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10c<- all10c[complete.cases(all10c),]
# 1106 cases - 13 vars

#make 3 new data sets from all9a instead of all9c. Ie long winded Business and JD.Second levels
all10d<- all9a %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol)
all10d<- all10d[complete.cases(all10d),]
# 2280 cases

all10e<- all9a %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10e<- all10e[complete.cases(all10e),]
# 1987 cases

all10f<- all9a %>% select(no.users, Discipline, pc.pro, Business, JD.Second,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10f<- all10f[complete.cases(all10f),]
# 1106 cases

```

First tune random forest parameters with caret package.
Try tuning with all 3 datasets

```{r, echo=FALSE}

# set seeds
set.seed(123)
seeds <- vector(mode = "list", length = 51)
for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
## For the last model:
seeds[[51]] <- sample.int(1000, 1)

d.sets <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")

# must turn outcome into two level factor
for( i in 1:length(d.sets)){
        df = get(d.sets[i])
        df$f.rpdol<- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol)=="0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol)=="1"] <- "loss"
        assign(d.sets[i], df, envir = .GlobalEnv)
}

#function for running caret..

caret.all <- function(df = all10rf){
        set.seed(100)
        sample<- sample(1:nrow(df), 2/3*nrow(df), replace=F)
        train<- df[sample,]
        test<- df[-sample,]

        if(nrow(df) == 2364){
                formula<- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type"

        } else{
                if(nrow(df) <= 2300){
                        formula<- "f.rpdol ~ . -Year - mlsto -b.rpdol"
                }
        }

        cv.Control <- trainControl(method = "cv",
                           number = 10,
                           seeds= seeds
                           , summaryFunction= twoClassSummary,
                           classProbs=TRUE
                           )
        gbmGrid <- expand.grid(mtry = c(3,4,5,6,7))

        set.seed(2)
        gbmFit <- train(as.formula(formula), data= train,
                method = "rf", trControl = cv.Control, verbose = FALSE,
                tuneGrid = gbmGrid,
                metric = 'ROC', allowParallel = T, ntree= 1000)

        plot(gbmFit)

}

caret.all(df = all10rf) # mtry = 4, AUC = 0.7545 
caret.all(df = all10mice) # mtry = 5, AUC = 0.757
caret.all(df = all10a) # mtry = 3, AUC = 0.719
caret.all(df = all10b) # mtry = 4, AUC = 0.771
caret.all(df = all10c) # mtry =4, AUC = 0.7735

# all10c is performing best! let's make a function that runs each df through the test set?
df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")

test.roc <- function(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "log", x.folds = 5,
                     extra.var = NULL, mult = 1){

        auc<- rep(list(rep(0, length(dfs))), length(dfs))
        colours <- topo.colors(x.folds*mult)

        for(i in 1:length(dfs)){
                df = get(dfs[i])

                # mlsto is spinning errors, get rid of
                df = df[, !(names(df) %in% 'mlsto')]
                set.seed(seed)

                # create f.rpdol
                df$f.rpdol <- as.factor(df$b.rpdol)
                levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
                levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

                #set up 10 fold cv using caret function
                folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                #now have folds[[1]] through folds[[10]] list of rows to exclude

                if(mult > 1){
                        for(k in 2:mult){
                                set.seed(sample(1:1000,1))
                                folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                                folds = append(folds, folds.temp)
                        }
                }

                #if extra.var not found in names(df), change to NULL
                if(!is.null(extra.var)){
                        if(!(extra.var %in% names(df))) extra.var = NULL
                }

                #define formula for algorith methods, either rf, log, boost

                if(nrow(df) == 2364){
                        if(is.null(extra.var)){
                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

                        } else {
                                formula<- paste("f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second", "+", extra.var, sep = " ")
                        }

                } else{
                if(nrow(df) <= 2300){
                        # include code.client, code.contact, Billing.Type
                        formula<- "f.rpdol ~ . -Year -b.rpdol "
                }
                }

                if(method %in% 'boost') {
                        #no JD.Second
                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

                }

                if(method %in% 'naiive') {
                        formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"

                }

                #for loop for each fold

                for (j in 1:length(folds)) {

                        #turn predictors and response into 'ordered' variables for roc() to work
                        test<- df[folds[[j]],]
                        levels(test$f.rpdol) <- c('profit', 'loss')
                        test$f.rpdol <- ordered(test$f.rpdol)

                        if(method %in% 'rf'){
                                fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                            mtry = mtrys[i], ntree = 1000)
                                pred <- predict(fit, test, type = "prob")
                                pred <- pred[,2]

                        }

                        if(method %in% 'log'){
                                fit<- glm(as.formula(formula), 
                                          data = df[-folds[[j]],], family = binomial())
                                pred <- predict(fit, test,
                                                type = "response")
                        } 
                        if(method %in% 'boost'){
                                fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
                                           distribution = "bernoulli", n.trees = 10000,
                                           shrinkage = 0.001, interaction.depth = 5,
                                           n.minobsinnode = 20)

                                pred <- predict(fit, test, n.tree = 10000, type = "response")
                        }

                        if(method %in% 'naiive'){
                                fit<- naiveBayes( as.formula(formula), df[-folds[[j]],])
                                pred <- predict(fit, test, type = 'raw')
                                pred<- pred[,2]
                        }

                        #pROC
                        pred.p <- pROC::roc(test$f.rpdol, pred)

                        if (i == 1 & j== 1) {
                        plot.roc(pred.p, col = colours[i], print.thres = F)

                        } else{
                        plot.roc(
                        pred.p, col = colours[j], add = TRUE , print.thres = F)
                        }

                        auc[[i]][j]<- auc(pred.p)
                        names(auc)[[i]] = method

                        #print j value
                        cat(j, " ")

                        }

                print(i)

        }

        assign("auc.func", auc, envir = .GlobalEnv)
        assign("formula", formula, envir = .GlobalEnv)

}

df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 400, mtrys = c(4,5,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# for random forest, the all10rf performs second/third best and most consistently, about AUC = 0.765
```

Now run logistic regression

```{r, echo= FALSE}

df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")
test.roc(dfs = df.list, seed = 400, mtrys = c(4), method = "log",
         x.folds = 5, extra.var = NULL, mult = 1)
names(auc.func) = df.list
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# logistic regression does well! with all10mice, AUC = 0.775

```

Try naiive bayes - try full NA dataset, all9c

```{r naiive bayes, echo=FALSE}

formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"

df = all10mice
df = all9c

set.seed(sample(1:1000,1))
sample<- sample(1:nrow(df), 2/3*nrow(df), replace=F)
train<- df[sample,]
test<- df[-sample,]

fit<- naiveBayes( as.formula(formula), train)
pred <- predict(fit, test, type = 'raw')
pred<- pred[,2]

pred.p <- roc(test$b.rpdol, pred)
plot.roc(pred.p, col = colours[i], print.thres = T)

#naiive bayes can do about 0.71 at best!! logistic regression, random forest and  boosted trees are
# definitely doing something!!!

```

Test if data sets subset from all9a (full original Business and JD.Second) instead of all9c perform better

```{r, echo=FALSE}

#create factor profit/loss variable
d.sets<- c("all10d", "all10e", "all10f")

for( i in 1:length(d.sets)){
        df = get(d.sets[i])
        df$f.rpdol<- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol)=="0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol)=="1"] <- "loss"
        assign(d.sets[i], df, envir = .GlobalEnv)
}

#perform 10 fold cv with same parameters as 10a, 10b, 10c
df.list <- c( "all10a", "all10b", "all10c", "all10d", "all10e", "all10f")
test.roc(dfs = df.list, seed = 100, mtrys = c(3,4,4,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#try log reg
test.roc(dfs = df.list, seed = 100, mtrys = c(3,4,4,3,4,4), method = "log")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")
#performed much better with compacted categories

```

Experiment without JD.Second:

```{r, echo=FALSE}
df.list <- c("all10rf", "all10mice")
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5), method = "log")
with.JDauc<- melt(auc.func)

#remove 'JD.Second' from formula and re-run test.roc
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5), method = "log")
witho.JDauc<- melt(auc.func)

all.JDauc<- rbind(with.JDauc, witho.JDauc)
all.JDauc[21:30,]$L1<- 3
all.JDauc[31:40,]$L1<- 4

ggplot(all.JDauc, aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
t.test(all.JDauc$value[1:10], all.JDauc$value[21:30])
t.test(all.JDauc$value[11:20], all.JDauc$value[31:40])

#JD.Second really doesn't matter, pretty much just noise, meh keep it in.

```

See what effect including clients has on random forest and log.
It did not help boosted trees.
Experiment on all10mice dataset.

Create five experimental datasets:

1. all10mice.c1: complete cases of all10mice once added code.client
2. all10mice.c2: only include clients with over 20 count. rest become 'unknown'
3. all10mice.c3: only include clients with over 40 count. rest become 'unknown'
4. all10mice.c4: only include clients with over 60 count. rest become 'unknown'
5. all10mice.c5: only include clients with over 70 count. rest become 'unknown'

```{r, echo=FALSE}
#create data frame of client codes and their counts, as well as contacts
client.count<- ddply(all10mice, .(code.client), nrow)
client.count<- arrange(client.count, -V1)
contact.count<- ddply(all10mice, .(code.contact), nrow)
contact.count<- arrange(contact.count, -V1)

#all10mice.c1
client.count2<- client.count[1:53,]
all10mice.c1<- all10mice %>% filter(code.client %in% client.count2$code.client)
all10mice.c1<- all10mice.c1 %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, JD.Second, majority.pos, pc.majpos.log,
                         Billing.Type, code.client, f.rpdol)
all10mice.c1 <- all10mice.c1[complete.cases(all10mice.c1),]
all10mice.c1$code.client<- droplevels((all10mice.c1$code.client))
all10mice.c1$reduc.client<- all10mice.c1$code.client

#four new data frames for different client thresholds

all10mice.c2 <- all10mice
all10mice.c3 <- all10mice
all10mice.c4 <- all10mice
all10mice.c5 <- all10mice

add.clientvar<- function(df = c('all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5'), 
                         thresh = c(20, 40, 60, 70), thresh.contact = c(5,8,13,20)) {

        for (i in 1:length(thresh)) {

                df.temp = get(df[i])
                threshold = thresh[i]

                #client code first
                client.count2 <- client.count[client.count$V1 > threshold &
                                                      !is.na(client.count$code.client),]

                df.temp$code.client <- as.character(df.temp$code.client)
                df.temp$reduc.client <- ifelse(df.temp$code.client %in% client.count2$code.client,
                                             df.temp$code.client,
                                             'unknown')
                df.temp$code.client <- as.factor(df.temp$code.client)
                df.temp$reduc.client <- as.factor(df.temp$reduc.client)

                #contact code second
                contact.count2 <- contact.count[contact.count$V1 > thresh.contact[i] &
                                                      !is.na(contact.count$code.contact),]

                df.temp$code.contact <- as.character(df.temp$code.contact)
                df.temp$reduc.contact <- ifelse(df.temp$code.contact %in% contact.count2$code.contact,
                                             df.temp$code.contact,
                                             'unknown')
                df.temp$code.contact <- as.factor(df.temp$code.contact)
                df.temp$reduc.contact <- as.factor(df.temp$reduc.contact)

                assign(df[i], df.temp, envir = .GlobalEnv)

                }
        }

add.clientvar(df = c('all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5'), 
                         thresh = c(20, 40, 60, 70), thresh.contact = c(5,8,13,20))

#now have 5 datasets to compare, all derived from client count and all10mice
# want to run each of these through 10 fold cross validation to see if we get better/worse results

#tune first
caret.all(df = all10mice.c1)

#10 fold test/train cv for code.client
df.list<- c('all10mice', 'all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5')
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "rf",
         extra.var = 'reduc.client')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "log",
         extra.var = 'reduc.client')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#10 fold test/train cv for code.contact
df.list<- c('all10mice', 'all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5')
test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "rf",
         extra.var = 'reduc.contact')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "log",
         extra.var = 'reduc.contact')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# including client.code really doesnt matter.
# including contact.code seems to make it marginally worse
# leave.. code.contact out, leave code.client in maybe.

#save all10mice.c4 for BN trial
write.csv(all10mice.c4, 'C:/Users/n9232371/Documents/Consultbusiness/data/all10mice_c4.csv')

```

Compare

* logistic regression
* random forest
* boosted tree

```{r, echo = FALSE}

all10mice.c4 <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice_c4.csv')
df.list<- c('all10mice')

#logistic regression with all10mice
test.roc(dfs = 'all10mice.c4', seed = 100, mtrys = c(5), method = "log",
         extra.var = NULL, mult = 3)
auc.log.noBil = auc.func
auc.log = auc.func
auc.log.nocli = auc.func

#random forest with all10mice
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "rf",
         extra.var = NULL, mult = 3)
auc.rf = auc.func

#boosted tree with all10mice
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "boost",
         extra.var = NULL, mult = 3)
auc.boost = auc.func

auc.func<- list(auc.log, auc.rf, auc.boost, auc.log.noBil, auc.log.nocli)

ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[61:90], auc$value[91:120])

# no significant difference. I'm leaning towards boosted tree.
```

Look at variable importance

```{r, echo= FALSE}
formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

#random forest
fit.rf <-randomForest(as.formula(formula), data = all10mice,
        mtry = 5, ntree = 1000, importance = T)
imp<- as.data.frame(importance(fit.rf))
imp<- imp[order(imp$MeanDecreaseAccuracy,decreasing=TRUE), ]

#logistic regression
fit.log <- glm(as.formula(formula),
           data = all10mice, family = binomial())

summary(fit.log)

#boosted tree
formula<- formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

fit.boost <-  gbm(as.formula(formula), data = all10mice,
        distribution = "bernoulli", n.trees = 10000,
        shrinkage = 0.001, interaction.depth = 4,
        n.minobsinnode = 20)

summary(fit.boost)
saveRDS(fit.boost, 'C:/Users/n9232371/Documents/Consultbusiness/data/fit_boost.rds')
fit.boost<- readRDS('C:/Users/n9232371/Documents/Consultbusiness/data/fit_boost.rds')

#partial dependence plots
fit.boost$var.names
plot.gbm(fit.boost, 3, type="response")
biz<- plot.gbm(fit.boost, 7, type="response")

##
### cforest
formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

system.time(
fit.cforest<- cforest(as.formula(formula),
        data= all10mice
        )
)

var.imp<- varimp(fit.cforest) %>% as.data.frame
colnames(var.imp)[names(var.imp) %in% '.']<-'imp'
var.imp$var <- rownames(var.imp)
#reorder levels for bar plot
var.imp <- within(var.imp, 
                    var <- factor(var,levels=
                                            var.imp[order(-abs(var.imp$imp)),]$var))

q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
a<- q + geom_bar(stat='identity')

ggsave(a, filename="cfor_all10mice.png", path=
               'C:/Users/n9232371/Documents/github/consult/finalwriteup/images/cforest_all10mice')

```

Look at results

```{r, echo = F}

#random forest
set.seed(200)
sample<- sample(1:nrow(all10mice), 2/3*nrow(all10mice), replace=F)
train<- all10mice[sample,]
test<- all10mice[-sample,]

formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

#random forest
fit.rf <-randomForest(as.formula(formula), data = train,
        mtry = 5, ntree = 1000, importance = T)
pred.rf <- predict(fit.rf, test, type = "prob")
imp<- as.data.frame(importance(fit.rf))
imp<- imp[order(imp$MeanDecreaseAccuracy,decreasing=TRUE), ]

#logistic regression
fit.log <- glm(as.formula(formula),
           data = train, family = binomial())
pred.log <- predict(fit.log, test,
                type = "response")
summary(fit.log)

#boosted tree
formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

fit.boost <-  gbm(as.formula(formula), data = train,
        distribution = "bernoulli", n.trees = 10000,
        shrinkage = 0.001, interaction.depth = 4,
        n.minobsinnode = 20)

pred.boost <- predict(fit.boost, test, n.tree = 10000, type = "response")
pred.p <- roc(test$b.rpdol, pred.boost)

plot.roc(pred.p, col = 'red', print.thres = T)
summary(fit.boost)

# look at test data and predictions
test.review <- test 
test.review$pred <- pred.boost 
test.review<- arrange(test.review, -pred)

write.csv(test.review, 'C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
```

Create plot for final report
Boosted tree

```{r}
df.list<- c('all9c')
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "boost",
         extra.var = NULL, mult = 1, x.folds = 10)
auc.func

ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

```

Create loop through each method - 20 models - compile AUC results

```{r}
test.roc(dfs = c('all10a'), seed = 100, mtrys = c(5), method = "log",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc = auc.func

test.roc(dfs = c('all9c'), seed = 100, mtrys = c(5), method = "boost",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc = append(tot.auc,auc.func)

test.roc(dfs = c('all10c'), seed = 100, mtrys = c(4), method = "rf",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc = append(tot.auc,auc.func)

test.roc(dfs = c('all9c'), seed = 100, mtrys = c(4), method = "naiive",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc = append(tot.auc,auc.func)

saveRDS(tot.auc,"~/Documents/github/consult/finalwriteup/report_data/reduc_df_auc.rds")

ggplot(melt(tot.auc), aes(x= factor(L1), y= value)) + geom_violin() + geom_jitter() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

# n.samples function is defined in 'draft.rmd'
index = c(1,3,4)
ns = NULL
for(i in seq_along(index)){
        temp = n.samples(data = tot.auc[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
        ns[i] = temp
}
ns

```

Re-run with MICE imputed data

```{r}

all10mice_f = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')

test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc.m = auc.func

test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "boost",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc.m = append(tot.auc.m,auc.func)

test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(4), method = "rf",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc.m = append(tot.auc.m,auc.func)

test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(4), method = "naiive",
         extra.var = NULL, mult = 10, x.folds = 10)
tot.auc.m = append(tot.auc.m,auc.func)

saveRDS(tot.auc.m,"~/Documents/github/consult/finalwriteup/report_data/mice_df_auc.rds")

ggplot(melt(tot.auc.m), aes(x= factor(L1), y= value)) + geom_violin() + geom_jitter() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

# n.samples function is defined in 'draft.rmd'
index = c(1,3,4)
ns = NULL
for(i in seq_along(index)){
        temp = n.samples(data = tot.auc.m[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
        ns[i] = temp
}
ns

```