### 2016-05-31 model_blend.rmd
started model blending

---
title: "Model Averaging"
author: "Amy Cook"
date: "May 26, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('caret')
library('pROC')
library('e1071')
library('randomForest')
library('party')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")

if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
}

```

Model averaging how to do it>?
Best models are ..

* boosted tree
* logisitic regression
* random forest

# For each method, create a complete dataset of predictions created from ten fold cross validation
```{r}

#create test and train set: 0.75/0.25
set.seed = 100
train = sample_frac(all10mice, 0.75, replace = FALSE)
test = setdiff(all10mice, train)

#perform 10-fold cross validation on train set to create complete train set of predictions for each method
full.predict <- function(df = train, method = "log", x.folds = 5, manual.seed = 100){

        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

        #set up 10 fold cv using caret function
        set.seed(manual.seed)
        folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
        #now have folds[[1]] through folds[[10]] list of rows to exclude

        #create NULL dataframe with row names from folds
        method.pred <- data.frame('rownames' = unlist(folds))
        temp.pred = NULL

        #define formula
        if(method %in% c('log','rf')){
                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
        }

        if(method == 'boost') {
                #no JD.Second
                formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
        }

        #loop over each fold, and compile predictions
        for (j in 1:length(folds)) {

                #turn predictors and response into 'ordered' variables for roc() to work
                test<- df[folds[[j]],]
                levels(test$f.rpdol) <- c('profit', 'loss')
                test$f.rpdol <- ordered(test$f.rpdol)

                if(method %in% 'rf'){
                        fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                    mtry = 4, ntree = 1000)
                        pred <- predict(fit, test, type = "prob")
                        pred <- pred[,2]

                }

                if(method %in% 'log'){
                        fit<- glm(as.formula(formula), 
                                  data = df[-folds[[j]],], family = binomial())
                        pred <- predict(fit, test,
                                        type = "response")
                } 
                if(method %in% 'boost'){
                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
                                   distribution = "bernoulli", n.trees = 10000,
                                   shrinkage = 0.001, interaction.depth = 5,
                                   n.minobsinnode = 20)

                        pred <- predict(fit, test, n.tree = 10000, type = "response")
                }

                temp.pred <- c(temp.pred, pred)

                #print j value
                cat(j, " ")

        }

        method.pred[, method] = temp.pred
        return(method.pred)

}

pred.log = full.predict(df = train, method = 'log', x.folds = 5, manual.seed = 100)
pred.rf = full.predict(df = train, method = 'rf', x.folds = 5, manual.seed = 100)
pred.boost = full.predict(df = train, method = 'boost', x.folds = 5, manual.seed = 100)

#merge all by rownames
pred.full = Reduce(full_join, list(pred.log, pred.rf, pred.boost))
gather.pred = gather(pred.full, key = method, value = pred.val,
                     -rownames)
order.rows = gather.pred %>% group_by(rownames) %>% 
        summarise(mean = mean(pred.val)) %>%
        arrange(mean) %>% ungroup
gather.pred$rownames <- factor(gather.pred$rownames,
                               levels= order.rows$rownames)

#plot spread of predictions

ggplot(gather.pred, aes(x = rownames, y = pred.val)) + geom_point(aes(colour = method))

ggplot(gather.pred, aes(x = method, y = pred.val)) + geom_boxplot(aes(fill = method), colour = "darkgray")

#randomForest seems to predict a bit higher than the other two methods. should I standardise it a bit??

```

Now create complete train set

```{r}
train.blend = full_join(train %>% mutate(rown = 1:nrow(train)), pred.full, 
                        by = c('rown'='rownames'))
train.blend <- train.blend %>% select(-rown)

```

# Try Feature weighted linear stacking

```{r cars}
#first write formula
# create f.rpdol
train.blend$f.rpdol <- as.factor(train.blend$b.rpdol)
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "0"] <- "profit"
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "1"] <- "loss"

formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + log + rf + boost"

#need to add interaction terms

```

# Try random forest

```{r pressure, echo=FALSE}
plot(pressure)
```

# Try boosted trees
```{r pressure, echo=FALSE}
plot(pressure)
```