### 2016-06-17 finalwriteup/trial_code.rmd
reviewed lit review started method review

---
title: "trials"
author: "Amy Cook"
date: "June 16, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$effort = \alpha*size^{\beta}$  
$\alpha = productivity \ coefficient$  
$\beta = economies \ of\ scale\ coefficient$  
$size = estimated \ lines \ of \ code$

$$
\begin{aligned}
 effort &= \alpha*size^{\beta} \\
 \alpha &= productivity \ coefficient \\
 \beta &= economies \ of\ scale\ coefficient \\
 size &= estimated \ lines \ of \ code
\end{aligned}
$$

* $project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
* $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$