### 2016-06-22 finalwriteup/trial_code.rmd
finished reviewing results to blended and added html

---
title: "trials"
author: "Amy Cook"
date: "June 16, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$effort = \alpha*size^{\beta}$  
$\alpha = productivity \ coefficient$  
$\beta = economies \ of\ scale\ coefficient$  
$size = estimated \ lines \ of \ code$

$$
\begin{aligned}
 effort &= \alpha*size^{\beta} \\
 \alpha &= productivity \ coefficient \\
 \beta &= economies \ of\ scale\ coefficient \\
 size &= estimated \ lines \ of \ code
\end{aligned}
$$

* $project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
* $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$

* total amount invoiced * percent of job by professional level employee
* total amount invoiced * client id
* project category * number of employees on the job

| Variable Description                                                 | ANOVA | RF | CF |
|----------------------------------------------------------------------|-------|----|----|
| % of Hours Completed by a Professional-level Employee                | x     | x  | x  |
| Timespan                                                             | x     | x  | x  |
| Number of Employees on Project                                       | x     | x  | x  |
| Amount Invoiced for Project                                          | x     | x  | x  |
| Business Category of Client                                          | x     | x  | x  |
| Project Category                                                     | x     |    | x  |
| Total Amount Invoiced from the Client - Past Jobs                    |       | x  |    |
| Discipline                                                           | x     |    |    |
| Position of Main Project Employee - new variable from ANOVA findings | x     |    |    |
| % of Hours by Main Project Employee                                  |       |    | x  |
| Billing Type (requested by case study company)                       |       |    |    |