### 2016-07-18 finalwriteup/trial_code_eq.rmd
started kerries changes

---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

```{r, echo=FALSE, include=FALSE}

library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('party')
library('randomForest')
library('RColorBrewer')
library('pwr')
library('scales')

# detect OS
if(.Platform$OS.type == 'windows'){
        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
} else{
        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
        # setwd("~/Documents/github/consult/finalwriteup/report_data")
        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
}

opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
```

$odds = e^{\ln\  odds}$

| Date    | Hours | Charge Amount (\$)| Cost Amount (\$)| Created By | Discipline | Job No. | Reconciled | Scope       |
|---------|-------|---------------|-------------|------------|------------|---------|------------|-------------|
| 2014/09/01 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
| 2014/09/03 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
| 2014/09/03 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
| 2014/09/08 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
| 2014/09/09 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
| 2014/09/15 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
| ...     | ...   | ...           | ...         | ...        | ...        |         | ...        | ...         |\
\\

$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$

$f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$

$RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$

where $X_{j}$ is simply something

$X_{1}$, include the terms $X_{2} = X_1^2$ and $X_{3} = X_1^3$ to build a third degree polynomial. Interactions between explanatory variables can be modelled by including a terms such as $X_5 = X_4*X_1$

$\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$

$$
\begin{aligned}
e^{\log(odds)} &= e^{\log\left(\frac{p}{1-p}\right)} \\ 
&= e^{-\log\left(\frac{1-p}{p}\right)} \\ 
&= e^{-\log\left(\frac{1}{p} - 1\right)} \\ 
&= 1 - \frac{1}{p}
\end{aligned}
$$