### 2016-12-13 SS/Regression.rmd
reset

---
title: "Regression"
author: "Amy Cook"
date: "August 15, 2015"
output: word_document
---

```{r, echo=FALSE, include=FALSE}

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')

library(devtools)
library('ggbiplot')
library('fpc')
library('vegan')
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d$Post.Code<- as.factor(all7d$Post.Code)

```

This file is to do a multi-step regression to predict RETURN PER DOLLAR.
The reason for multistepping is to allow for multiple variables with lots of missing NA values.

Variables to include:

Full:

inv.mlsto
timespan
no.users
Discipline
pc.pro
client.totinv
code.director
Num.disc
Business
Biz.type
code.ProjEng

Many NA:

code.contact
code.client
Project.Value
JD.Second
Post.Code
Billing.Type
majority.pos
pc.majpos

The 'Many NA' variables will be split into groups:

* code.contact: 1652 NA
* code.client: 422 NA
* Project.Value: 2333 NA
* JD.Second: 1057 NA
* Post.Code: 356 NA
* Billing.Type: 1314 NA
* majority.pos: 377 NA
* pc.majpos: 377 NA

```{r echo=FALSE}
some <- all7d %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
                         client.totinv,code.director,Num.disc,Business,Biz.type,
                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
                         pc.majpos)

#heatmap of any similar NA value ranges

experiment<- sapply(some, function(x) is.na(x))
NA.plot<- melt(experiment)
NA.plot$mlsto<- rep(all7d$mlsto, ncol(experiment))
#reorder levels for ggplot by total number of NA TRUE
# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
NA.plot <- within(NA.plot, 
                         Var2 <- factor(Var2, 
                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
                                                                 decreasing=FALSE))))

ggplot(NA.plot, aes(x=Var2, y=mlsto)) +
        geom_tile(aes(fill = value)) + 
        theme(axis.text.x=element_text(angle=45,hjust=1))

```

Will address each of the variables with many nA separately. 8 of these variables totally. run a separate lm for each of these

Need to transform some numeric variables:

Full:

return.pdol - cube root
inv.mlsto - log
timespan - cube root
pc.pro - leave alone
client.totinv - log

Many NA:

Project.Value - log
pc.majpos - log

```{r, echo=FALSE}

#write a function for testing transformations

trans<- function(var= 'timespan', df= all7d, transform = 'cbrt'){
        #initial q plot
        init<- qqPlot(df[,var], 
                       main = 'untouched variable')
        print(init)

        #make trial dataframe
        trial = sapply(df[,var], function(x){
                if(transform == 'log'){
                        log(x)
                } else

                if(transform == 'neglog')
                        {
                        sign(x)*log(abs(x) + 1)
                } else

                if(transform == 'sqrt')
                {
                        sqrt(x)
                } else
                        if(transform == 'cbrt')
                {
                        sign(x) * abs(x)^(1/3)
                } else
                { x^2 }

                }) %>% 
                as.data.frame

        colnames(trial)[names(trial) %in% '.']<-'trans.var'

        #qplot for transformation
        qqtrans<- qqPlot(trial$trans.var, 
                         main = 'transformed')
        print(qqtrans)

        ##move on to density plots

        #original variable
        dens<- ggplot(df, aes_string(x=var)) + geom_histogram(aes(y=..density..))
        print(dens)

        #transformed variable
        dens.trans<- ggplot(trial, aes_string(x='trans.var')) + geom_histogram(aes(y=..density..))
        print(dens.trans)
}

trans(var= 'timespan', df= all7d, transform = 'cbrt')
trans(var= 'inv.mlsto', df= all7d, transform = 'log')
trans(var= 'return.pdol', df= all7d, transform = 'cbrt')
trans(var= 'pc.majpos', df= all7d, transform = 'log')
trans(var= 'client.totinv', df= all7d, transform = 'log')
trans(var= 'Project.Value', df= all7d, transform = 'log')

```

Have now decided on transformations for numeric variables.

Create all7d ready for linear regression analysis

```{r, echo=FALSE}

math.cbrt <- function(x) {
        sign(x) * abs(x)^(1/3)
}

all7d$timespan.cbrt<- math.cbrt(all7d$timespan)
all7d$inv.mlsto.log <- log(all7d$inv.mlsto)
all7d$client.totinv.log <- log(all7d$client.totinv)
all7d$Project.Value.log <- log(all7d$Project.Value)
all7d$pc.majpos.log <- log(all7d$pc.majpos)

```

Root mean square error as initial check. Write code which transforms response variable back to original format by cubing, then finds RMSE.

```{r, echo=FALSE}

#predict Petal.Width using lm, find RMSE
#answer will be return.pdol
#guess will be response of prediction which is cube root
RMSE<- function(answer, guess){
        error= answer-guess
        root.error= sqrt(mean(error^2))
        print(root.error)
}

```

Now fit full variables to linear regression

```{r, echo=FALSE}

full = lm(return.pdol ~ 
                  inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + client.totinv.log + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng,
          data = all7d)
summary(full)

#have a look at the residuals

res<- resid(full)
plot(all7d$inv.mlsto.log, res)
abline(0,0)

plot(density(res))
qqPlot(res)

#compare some of the answers
ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full, all7d, interval= 'predict'))

RMSE(ans$answer, ans$full.lm)

```

RMSE is 0.53 which isn't great in terms of predicting return per dollar.
R^2 is 0.2152

Let's try adding more variables:

```{r, echo=FALSE}
#non full variables
#Post.Code
PostCode<- lm(return.pdol.cbrt ~ Post.Code, data=all7d)
summary(PostCode)
plot(Post.Code ~ return.pdol.cbrt, data=all7d)
abline(PostCode)

#write function

col.lm<- function(vars= 'Post.Code', df=all7d, ret= 'return.pdol.cbrt'){
        #loop through and create lm model's
        names= NULL

        for(i in 1:length(vars)){
                formula = paste(ret, "~", vars[i], sep = " ")
                linear = lm(as.formula(formula), data = df)
                assign(paste(vars[i],"lm",sep="."), linear, envir=.GlobalEnv)

                names = rbind(names, paste(vars[i],"lm",sep="."))

                cat(i," ")
        }

        names = names %>% as.vector
}

#now put all incomplete variables through col.lm function
varis= c('Post.Code', 'majority.pos','pc.majpos.log','code.client', 'JD.Second',
        'Billing.Type','code.contact','Project.Value.log')

lm.vec<- col.lm(vars= varis, df=all7d, ret= 'return.pdol.cbrt')

#compile results of all lm's into ans dataframe

ans.complete<- function(lms = 'Post.Code.lm', df=all7d, ans.df= ans){
        ans.df=ans
        for(i in 1:length(lms)){
                column = predict(get(lms[i]), df)
                #transform from cube root
                column = column^3
                ans.df = cbind(ans.df, column)

                #rename column
                colnames(ans.df)[i+2]<-lms[i]

                cat(i," ")
        }

        ans.df
}

ans.full<- ans.complete(lms=lm.vec, df=all7d, ans.df=ans)

# RMSE(ans$answer, ans$majority.pos.lm) --  doesn't work, need to adjust RMSE function

```

So now we have our predictions for the full variables as well as a column for each of the incomplete variables.

We want to recombine them using linear regression again!

For a check, I will find all cases which have the full variables as well as client.code, code.contact, and Billing.Type

```{r, echo=FALSE}
#reduce ans data frame to only complete cases with variables you are interested in
ans.red = ans.full %>% select(answer, full.lm, code.client.lm, Billing.Type.lm, code.contact.lm)
ans.red <- ans.red[complete.cases(ans.red),]

#run final lm
final.lm<- lm(answer~. , data=ans.red)
final<- predict(final.lm, ans.red)
ans.red$final.lm<- final

RMSE(ans.red$answer, ans.red$final.lm)

ans.red %>% View

```

Wouldnt you be better off just running a regression with the complete cases?

```{r, echo=FALSE}

reduced<- all7d %>% select(inv.mlsto.log , timespan.cbrt , no.users , Discipline , pc.pro , client.totinv.log , 
                                   code.director , Num.disc , Business , Biz.type , code.ProjEng, Billing.Type,
                           code.client, code.contact, return.pdol.cbrt)
reduced<- reduced[complete.cases(reduced),]

test.lm<- lm(return.pdol.cbrt ~ ., data= reduced)
test<- predict(test.lm, reduced)
ans.red$test.lm <- test

RMSE(ans.red$answer, ans.red$test.lm)

```