### 2016-12-13 prob_review.rmd

---
title: "result_review.rmd"
author: "Amy Cook"
date: "November 19, 2015"
output: word_document
---

```{r_loadpackages, echo=FALSE, include=FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.3/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.3/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.3/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.3/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.3/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.3/library')
library('pROC',lib='C:/Progra~1/R/R-3.2.3/library')

# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

Load data

```{r, echo=FALSE}

all7d <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')

```

Assemble full data set with predictions

Write function for making 10 boosted trees and checking the predictions/savings for each!

```{r, echo=FALSE}

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]

predict.all<- function(df.string = "all10mice", method = "boost", seed = 300, k.folds = 3){

        #create dataset for final results to tag onto
        all10full = NULL

        df = get(df.string)
        set.seed(seed)

        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <-"profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <-"loss"

        #order levels
        df <- within(df, b.timespan.cbrt <- factor(b.timespan.cbrt, 
                                               levels=c('1d-3wk', '3wk-2.5m','2.5m-9m','9m-1.5y','1.5y-3y','>3y')))
        df <- within(df, majority.pos <- factor(majority.pos, 
                                               levels=c('contracttech', 'midtech','seniortech','gradpro',
                                                        'midpro','seniorpro','director')))

        levels(df$b.timespan.cbrt) <- c('1d-3wk', '3wk-2.5m','2.5m-9m','9m-1.5y','1.5y-3y','>3y')
        df$b.timespan.cbrt <- ordered(df$b.timespan.cbrt)

        #set up 10 fold cv using caret function
        folds <- createFolds(df$f.rpdol, k = k.folds, list = T, returnTrain = F)
        #now have folds[[1]] through folds[[10]] list of rows to exclude

        if(method %in% 'boost') {
                        formula = "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
        } else {
                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
        }

        for (j in 1:length(folds)) {

                        #turn predictors and response into 'ordered' variables for roc() to work
                        test<- df[folds[[j]],]
                        levels(test$f.rpdol) <- c('profit', 'loss')
                        test$f.rpdol <- ordered(test$f.rpdol)

                        if(method %in% 'rf'){
                                fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                            mtry = 5, ntree = 2000)
                                pred <- predict(fit, test, type = "prob")
                                pred <- pred[,2]

                        } else {
                                if(method %in% 'log'){
                                        fit<- glm(as.formula(formula), 
                                                  data = df[-folds[[j]],], family = binomial())
                                        pred <- predict(fit, test,
                                                        type = "response")
                                } else {
                                        fit <- gbm(as.formula(formula), data = df, #[-folds[[j]],],
                                                   distribution = "bernoulli", n.trees = 10000,
                                                   shrinkage = 0.001, interaction.depth = 4,
                                                   n.minobsinnode = 20)

                                                pred <-predict(fit, test, 
                                                               n.tree = 10000, type = "response")
                                                }
                        }

                        #pROC
                        pred.p <- roc(test$f.rpdol, pred)
                        # plot.roc(pred.p, col= colours[i], print.thres=T)

                        auc(pred.p)

                        ##
                        #create full test df
                        test$pred<- pred
                        test.pred <- merge(test, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")

                        all10full <- rbind(all10full, test.pred)

                        #print j value
                        cat(j, " ")

        }

        return(all10full)

}

results.1 <- predict.all(df.string = "all10mice", method = "boost", seed = 400, k.folds = 8)

#partial dependence plot for boosted only
fit$var.names
i=5
a<- plot.gbm(fit, i, type="response")
a

```