### 2016-12-13 summit_report/summit_report.rmd

---
title: "BT_insights"
author: "Amy Cook"
date: "October 7, 2016"
output:
  word_document: default
  html_document: default
---

```{r loadpackages, echo=FALSE, include=FALSE}
library('knitr')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('randomForest')
library('caret')
library('pROC')
library('testthat')
library('purrr')
library(tidyr)
library(broom)

opts_chunk$set(fig.width=6.7, fig.height = 5.5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

```{r, include = FALSE}
if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("C:/Users/n9232371/Documents/github/consult/functions.R")
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("/Users/yam/Documents/github/consult/functions.R")

}

```

The output below plots the tendency for a project to be loss making (y-axis) against an explanatory variable (x-axis).

The further up the y-axis a point sits, the more likely the project is to be loss-making is the simplest way of thinking about it.

On the x-axis, some of the category labels are missing, so I've printed out all the categories in the line below.

Also, some of the variable names on the x-axis are complicated, so I've added a comment after each plot which will help you interpret the x axis.

```{r, cache = TRUE, }

formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

fit <- gbm(as.formula(formula), data = all10mice,
                                   distribution = "bernoulli", n.trees = 10000,
                                   shrinkage = 0.001, interaction.depth = 5,
                                   n.minobsinnode = 20)

comments <- c(
        Discipline = "striaghtforward",
        pc.pro = "% of hours completed by an engineer on a project",
        b.timespan.cbrt = "timespan category - timespan of a project from first hour logged til last",
        no.users = "number of employees that worked on a project",
        b.inv.log = "total invoiced amount for a project ($) - ignore x-axis title",
        client.totinv.log = "average amount invoiced in total for a client across all jobs with them. 
        You need to take the natural exponent of the numbers on the x axis (i.e. e^6 = $403)",
        Business = "business category of the client, i.e. architect",
        majority.pos = "position of employee that completed the majority number of hours on a project.
        'tech' refers to drafters, 'pro' refers to engineers",
        pc.majpos.log = "% of hours completed by the main employee on the project. You need to take the natural 
        exponent of these values on the x-axis, i.e. e^4 = 55%",
        JD.second = "category of the project, i.e. subdivisions",
        Billing.Type = 'self explanatory'

)

# partial dependency plots
for(i in 1:11){
        plot.gbm(fit, i, type="response")
        print(fit$var.levels[[i]])
        print(paste("\n"))
        print(comments[i])
}

```