### 2016-12-19 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
finished results and discussion of paper

---
title: Predictive Models to Support Quoting of Fixed Fee Consulting Projects
author:
  - name: Amy Cook, Paul Wu, Kerrie Mengersen
    email: a21.cook@qut.edu.au
    affiliation: Queensland University of Technology
    footnote: Corresponding Author
address:
  - code: Queensland University of Technology
    address: School of Mathematical Sciences, George Street, Brisbane, QLD, 4000
abstract: |
  A concise abstract is required. limit to 250 words. clearly state purpose of research, principal resutls and major conclusions. no refs

  Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts. It was found that in a specific decision framework, a case study's profits could be improved 9% by declining approximately 4% of projects. Alternative decision frameworks are also proposed and evaluated. Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods. A decision scenario which rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements. The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements. The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.
keywords: |
  consulting; machine learning; profitability; predictive model; construction industry; data mining
abbreviations: 
  - name: Customer Relationship Management
    footnote: Abbreviations

bibliography: Library.bib
output: rticles::elsevier_article
---

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

```{r, echo=FALSE, include=FALSE}

library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('randomForest')
library('RColorBrewer')
library('pwr')
library('scales')
library('purrr')

# detect OS
if(.Platform$OS.type == 'windows'){
        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
} else{
        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
        # setwd("~/Documents/github/consult/finalwriteup/report_data")
        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
}

opts_chunk$set(fig.pos = 'h', dpi = 200, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)

# setwd("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
```

1. Introduction
==========================

Clearly state the research question and objectives of the work. Briefly provide any necessary
background to frame the research question. Concisely summarize the major findings/results.

Many consulting projects can be financialy risky endeavors for the consultant, often depending on the industry. Engineering consultanting companies in the construction industry complete a wide range of complex projects in a competitive enviroment. Because of the way contracts are structured, many of these projects result in losses. The purpose of this study was to test whether these loss making projects could be predicted by statistical and machine learning algorithms using the information available at project conception. Secondly, could the predictive algorithm improve a company's profitability? Predictive models and value frameworks were built using project data from a case study engineering company to answer these questions.

1 page

#### 1.1 Problem motivation

Financial risk is taken on by a consulting company when they offer a fixed fee to complete a project. In competitive climates, clients are in the position to request fixed price quotes from several consultants before a project is awarded. Hence, initial quoted fees must be competitive but also accurately cover consultant salaries and business costs.

How do consultants make this accurate yet competitive fee estimate? This varies from industry to industry, but typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, analysis of project details, and rules of thumb. Unfortunately, some projects priced using this method can blow out of budget, causing significant financial losses to the consultant.

Cost blowouts of complex projects has a history, particularly documented in the It and construction industry. A review of 6 IT project surveys between 1984 and 1994 by Moløkken & Jørgensen (2003) revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by an average of 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].

Kahneman won a nobel prize in 2002 for his research into decision making on this topic. He argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project [@Lovallo2003]. To counter this phenomenon, research by Flyvbjerg [-Flyvbjerg2011] advised that managers should be confronted by the financial performance of similar projects to the one being costed so that they take on a more realistic view of their abilities. This was based off a body of findings where people made better predictions about themselves after being exposed to facts about other people. For example, subjects were asked to predict their skill levels before and after being exposed to a summary of other people's skills, and their predictions after were significantly more accurate [@Lovallo2003].

This work aims to help managers provide better fixed price quotes by predicting whether a new project will be profitable or not, based on an algorithm trained on all previous project data from the same company. The manager could take multiple actions based on the algorithms output including not tendering for a project predicted to be loss-making, providing a high fixed price, or restructuring the contract so that it is not 100% fixed price. This research tests the simple case of rejecting all projects predicted as loss-making.

#### 1.2 Case Study

Project data from a single case study consulting company in the construction industry was used to build the predictive algorithm, trial several methods, and test their impact on the company's bottom line. Twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details was scraped from the company's internal database. Customer relationship management (CRM) software provided accesses to the database to all employees. Employees completed daily timesheets, alotting their hours to certian projects. Project managers input technical information and client details for each project, while the finance team tracked invoicing and additional project costs such as printing and transport in the same system. On top of this information, the CRM had the hourly cost of each level of employee. The cleaned dataset consisted of 2364 projects between 2004 and 2012. 70% of projects were fixed price and 20% of all projects made a loss after taking into account business costs.

Once several methods were built, to assess whether the model would improve the case study company's profitability, a simple managerial decision framework was applied: if the algorithm predicted the project to be loss making, the project was considered 'rejected' and all profits or *losses* of rejected projects were discarded. The remaining projects' profits and losses were summed to give a revised overall profit over the last 12 years of projects. This was compared to the summed profits and losses from all projects completed over the last 12 years. The best algorithm rejected enough loss-making projects to improve overall profits by 9%.

<!-- **General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data. Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.  -->
<!-- The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined. -->

<!-- **Hypothesis 1** -->
<!-- A statistical or machine learning model based on historical project data can predict the profitability of a new project with greater accuracy than a baseline predictor. A baseline predictor for this project is one that predicts the average of a numeric response variable for all cases or, if the response variable is categorical, randomly assigned categories for each case, where the proportion of assigned categories matches the true categorical proportions. -->

<!-- **Hypothesis 2** -->
<!-- The predictive model built from Hypothesis 1 can be shown to have a positive impact on the overall profit earned by the case study business. The overall profit is represented by the following equation: -->

2. Literature Review
====================

<!-- Summary of Key Related Research  -->
<!-- This section should include a brief summary of key related research. Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the goal is to  -->
<!-- clearly delineate a gap or missing link that the current  -->
<!-- research fills. Authors should avoid presenting a litany of past research and should focus -->
<!-- on prior work necessary to demonstrate the -->
<!-- existence of the research gap addressed in the manuscript. -->

For the case study company, training the algorithm to predict actual fee was not possible because the projects were too varied, and there was inadequate information describing the size and scope of each project in the database. However, in the literature, similar studies typically did attempmt to predict project costs of IT consutling projects and final construction costs of buildings and infrastructure.

#### 2.1 Cost estimation in the Construction Industry and IT Industry

Previous studies in the literature have focussed on reporting the predictive accuracy of their cost estimation models for IT or construction projects using data from multiple businesses or governments around the world. In contrast, this research is focussed on determining how or if models predicting project profitability can improve the bottom line for a case study business, given data the company has internally generated.

Reviewing a range of individual studies show that cost estimation models have generally been built from at most 530 projects [@Kim2004] [@Finnie1997] - 299 software development[@Pai2013] - 163 software [@Shin2015] - 234 building projects, and often under 100 projects [@Finnie1997] [@Chan2005] 87 buildings. Sometimes it is not stated where the project data came from, and some studies specified the information came from surveys, with responses from companies and governments around the world. In this study, the dataset originates from a single company's database, which means the data is easily accessible to the company and there are 2364 projects, which is much more than previous studies. A study by @Mendes2004 demonstrated that using within company data was also more accurate than across company data. Their dataset consisted of 67 across company web projects and 14 within company projects.

Almost all past studies analysed their model performance by predictive accuracy (reported as R^2 or RMSE) of cost predictions [@llist some stff] and compared these statistics bewteen methods. However, these results were not translated into a business case for the impact these models could have on an institution or how the model could be practically integrated into decision making. For example, if an institution used the model, how would it be applied to their decision making process and what magnitude of benefits could they expect? Saradhi & Palshikar's [-@Saradhi2011] work on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time, presented a good framework for implementing their model. Theirsupport vector machine (SVM) predictive model identified employees at high risk of churn. Then a method for determining the value of each employee in terms of their importance to projects and their monthly chargeability was described, which enabled high risk employees to be ranked. Managers then had a clear direction for who to target first to prevent churn. This extension provided a comprehensive framework for how business managers could adopt a machine learning algorithm to improve business operations. It was a valuable addition that is absent from most cost estimation research, but will be included in this work.

#### 2.2 Methods used in Cost Estimation and Other Business Applications

Cost estimation literature tended to apply linear regression, neural networks and sometimes SVM methods, while predictive methods applied to other business problems investigated a wider range. This study aims to compare a wider range of methods to past cost estimation studies.

In the construction industry, often only one method was assessed, such as linear regression, without comparison to other methods [@Chan2005; @Elfaki2014]. Neural Networks started appearing in literature in the 1990's and @Elfaki2014's review of cost estimation methods from 2004 to 2014 found that artificial Neural Networks and SVM's were the most popular machine learning techniques, but studies tended to only test one method. Some studies showed that Neural Networks outperformed Linear Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003]. To the authors' knowledge @Shin2015 documented the only application of ensemble tree methods (Boosted Trees) to cost estimation in construction projects. They found Boosted Trees slightly outperformed Neural Networks, but not significanatly.

In the software cost estimation literature, Linear Regression was the most popular method, however multiple studies showed that Neural Networks definitively outperform regression models  [@Finnie1997; @Pai2013; @Matson1993]. Despite the higher performance, Neural Networks were criticised for not providing reasoning or structure behind their predictions (black-box predictor) [@Finnie1997].

The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees, however research on other business problems utilised a wider range of methods. In financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification. In addition, ensemble trees such as Random Forests and Boosted Trees can output partial dependency plots and variable importance measures that provide insight into the model's calculations. This is a valuable asset oer black box predictors. @Kumar2007 performed a detailed review of statistical and machine learning techniques applied over 37 years in the context of bankruptcy prediction in banks and reported that although SVM's performed well, they are often complex and slow, requiring a great deal of memory. They also assessed method blending techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods such as Linear Regression or Boosted Trees alone.

To date, Linear Regression, Neural Networks, and Support Vector Machines (SVM's) have been comprehensively applied to cost estimation. This research does not test black-box predictors (Neural Networks and SVM's) because insight into the model's predictions is crucial for facilitating user uptake in the case study company. Ensemble tree methods such as Random Forests and Boosted Trees will be used instead, as they can perform as well as Neural Networks while providing insight to their output [@Caruana2006]. Blending of multiple predictive models is also applied in this work and has not yet been done in cost estimation.

#### 2.3 Gaps

In comparison to previous studies, this project advances the body of work on cost estimation in two ways. First, it focusses on building a model that is practical and profitable for a business to implement by using the case study's interenal database as well as testing a managerial decision framework based off the model output. This framework presents a clear measure of expected benefits provided by adopting the profitability model into a business' decision process. Secondly, advanced predictive methods are trialled which have been minimally tested on this problem. These include ensemble decision tree methods, such as Random Forests and Boosted Trees, and blends of multiple models built from different methods.

3. Prediction Methods
=====================

Should provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant method modifications should be described.

A range of simple and complex, linear and non-linear methods were applied to the problem including Logistic Regression, Random Forests, Gradient Boosted Trees, and Naiive Bayes. Furthermore, logistic regression, random forests, boosted trees and simple averaging were used to blend predictions from individual models. A brief description of each method is provided below.

<!-- Neural Netowrks and SVM were not purusued because black box -->
<!-- A previously mentioned disadvantage is that the NN algorithm is a black box because the internal structure is too complex for interpretation. They also require a lot of training data relative to other methods. -->

<!-- Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008]. Because it is so important to engage decision makers with a model that can explain its results, these two methods are not appropriate for the effort estimation problem.  -->

#### 3.1 Predictive methods

The aim of this research is to build a model that predicts whether a consulting project will be profitable or not. Therefore each method must be capable of binary classifcation. Binary classification output from a model is the probability, between 0 and 1, of an event occuring. In this case, the event is a project being loss-making.

##### 3.1.1 Logistic Regression

Logisitic regression models a linear relationship bewteen the log odds of a binary response variable and each of the explanatory variables. This is performed similarly to linear regression, but is transformed to probability of an event occuring by taking the inverse log of the log odds and rearranging [@Macdonald1975]:

$$
\begin{aligned}
\log(odds) &= \beta_0 + X_1\beta_1 \\
\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 \\
-\log\left(\frac{1-p}{p}\right) &= \beta_0 + X_1\beta_1 \\
-\log\left(\frac{1}{p} - 1\right) &= \beta_0 + X_1\beta_1 \\
\log\left(\frac{1}{p} - 1\right) &= -\beta_0 - X_1\beta_1 \\
\exp\left(\log\left(\frac{1}{p} - 1\right)\right) &= \exp(-\beta_0 - X_1\beta_1) \\
\frac{1}{p} - 1 &= \exp(-\beta_0 - X_1\beta_1) \\
\frac{1}{p} &= 1 + \exp(-\beta_0 - X_1\beta_1) \\
p &= \frac{1}{1 + \exp(-\beta_0 - X_1\beta_1)}
\end{aligned}
$$

The result in a sigmoidal function bound by 0 and 1. A coefficient represents the change in the log odds of the response variable for each unit increase in an explanatory variable. Logistic Regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed [@Moore1989]. Addittionally, because of the simple, single linear relationship, the method is not prone to overfitting [@Perlich2003].

##### 3.1.2 Random Forests

Ensemble tree methods were pioneered in the 1990's to combat the low predictive accuracy and instability of individual decision trees. Combining hundreds or thousands of trees transform decision trees into high performing, stable predictors [@Breiman1996].

The Random Forests algorithm works by training multiple trees from the bootstrapped samples of a dataset. Additionally, when creating each tree, a random subset of attributes (covariates) is considered at each split. The reduced subset of covariates is resampled for each split in the tree. Dominant covariates are then suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker covariates [@Breiman2001a]. Performance statistics and tests run within the algorithm provide insights to the modeller, such as variable importance and variable relationships [@Sealfon2012].

##### 3.1.3 Gradient Boosted Trees

The boosted decision tree approach applies gradient descent theory to a series of decision trees. Performance tends to be higher if individual trees are limited to a certain depth to maintain simplicity and prevent overfitting [@Elith2008]. Unlike Random Forests, the trees are sequential, where each tree models the residuals (or errors) of the preceding tree.

By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the next tree more [@Hastie2009]. This increased weighting of error cases is called boosting. The successive results of up to thousands of trees is very powerful [@Elith2008]. @Caruana2006 tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests. This demonstrates how capable ensemble tree methods are of competing with high-level machine learning algorithms.

##### 3.1.4 Naive Bayes

Naive Bayes is a simple, fast method included in this work because, similarly to Logistic Regression, it provides a good baseline method against which to compare complex methods [@Caruana2006]. It works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for a categorical response variable. For example, if predicting a binary variable, $R$ (with classes 0 or 1) given two covariates $X_1$ and $X_2$, the equations for the probability of both outcomes is as follows:

$p(R = 1|X_1 X_2) = p(X_1|R = 1) \cdot p(X_2|R = 1) \cdot p(R = 1)$

$p(R = 0|X_1 X_2) = p(X_1|R = 0) \cdot p(X_2|R = 0) \cdot p(R = 0)$

These equations are simple to compute given the data, and the class with the highest probability can then be chosen [@Provost2013]. The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions, but amplifies the magnitude of the probability [@Provost2013]. This is fine for ranking class 0 against 1 but the output probabilities are not accurate statistical probabilities [@Caruana2006].

##### 3.1.5 Model Blending

Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009]. In this paper, the term 'blending' will be used to avoid confusion with ensemble tree methods. Model blending combines predictions from several models that used methods with different theoretical foundations. In this way, the strengths of each method can be combined. Since the 1990's research in statistics has shown that averaging results from different methods can provide better predictive accuracy than any single model [@Madigan1994].

Six blending methods, ranging from simple to complex, were chosen. These included simple averaging of the individual model results, building a Logistic Regression model and Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees. The first two models are simply the average, and weighted average of predictions from separate individual models. For example, if the Logistic Regression model output a probability of 0.6 for one case, while a Boosted Tree output 0.7, a simple blended average for that case would be a probability of 0.65. The last three methods are more complex, and take advantage of potentially meaningful interactions between the individual model predictions and the original covariates describing each case (called meta-features in the model blending literature) [@Sill2009]. For example, if the Logistic Regression model predicted profitability better than the Boosted Tree model for projects with only one team member, the complex blending methods would take advantage of this interaction. In summary, in the complex methods, the probability outputs from each model are interacted with each of the original covariates. Model blending has the potential to elevate output from the original methods by combining each of their strengths.

#### 3.2 Procedure

The raw data from the case study company's database consisted of three separate datasets: one describing the client and technical details for each project, another with each employee's time sheet entries, and another with all past invoices. This data was combined, cleaned, and rearranged so that each project was described within one row. From the timesheet data, 10 variables were engineered including the percent of hours performed by 'professional' employees as opposed to 'technical' employees, and the position of the employee that completed the most hours on each project. 3 variables were engineered from the invoicing data such as total amount invoiced for a project and mean invoice size for a client. Finally, from the project dataset, 2 variables were engineered including the number of projects completed with a client, and project category. Text analysis of the project description determined the project category variable. The response variable, whether the project was profitable or not, was calculated by subtracting the cost of employee hours from the timesheets from the sum of the paid invoices for a project. The total amount invoiced for projects ranged from \$500 to over \$1,000,000.

34 covariates made up the full dataset, and these were narrowed down to 11 in order to reduce predictive noise and provide a simple model that is comprehendable for stakeholders [@Weisberg2005]. Variable selection methods included linear regression, random forests, and cforests and the final variables are listed below.

* % of Hours Completed by a Professional-level Employee
* Time span
* Number of Employees on Project
* Amount Invoiced for Project
* Business Category of Client
* Project Category
* Total Amount Invoiced from the Client - Past Jobs
* Discipline
* Position of Main Project Employee - new variable from ANOVA findings
* % of Hours by Main Project Employee
* Billing Type

Timespan and the total amount invoiced are unknowns at before a project commences. Therefore, these variables were discretised into 5 or 6 categories from which a manager would be able to select. The category boundaries were determined by visualising clusters using hierarchical dendrograms.

##### 3.2.1 Research Question 1

To answer the first research question, if loss making projects can be predicted using statistical and machine learning algorithms, the binary classification and blending methods were built and evaluated. The area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to compare models and assess performance. Models were built on a training subset of data (80%) and validated on the remaining data. Therefore, different training and testing sets produced different AUC statistics. In order to compare which models performed significantly better than others, an adequate sample size of AUC statistics was required. A two-sample power calculation run on 20 AUC output for each model determined that 100 samples were required to achieve a statistical power of 0.8.

Then, the model output from the validation sets were combined and fed into the three simple and three complex blending models. Again, 100 models were built to achieve a power of 0.8 for the variation in AUC test statistics across different divisions of training/testing data. A maximum of five training and validation divisions were created from a complete set of individual model results.

##### 3.2.2 Research Question 2

The second research question asks whether a predictive algorithm can improve a company's profitability. The model with the highest AUC score may not improve company profits the most, because the magnitude of the profits and losses from each project have not been analysed. Therefore, all blended and individual models were carried forward to be assessed in a value framework.

Once a model outputs scores for the probability of each job being loss-making, the question then arises at what probability would a manager round the probability up to 1? And then what action would be taken? A threshold point between 0 and 1 can dictate when a project is rounded up, and classified as 'loss-making'. For this analysis, a simple decision scenario is tested where a manager rejects all projects with output above the threshold. For example, if the threshold is 0.6, and a project is predicted to have a probability of loss-making of 0.7, the project is rejected and profits or losses associated with the project are ignored.

To test each threshold between 0 and 1, profit curves were built for each method. This chart plots the probability threshold on the x-axis and the ratio of model-influenced profits over historical  company profits on the y-axis. The ratio represented by the y-axis is calculated using the formula below:

$$
\begin{aligned}
Ratio\ of\ Model\ Profit\ to\ Historical\ Profit\ (\%) &= \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
\end{aligned}
$$

<!-- \Delta\ Overall\ Profit\ \textrm{(%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100 -->

Where

$$
\begin{aligned}
I(\cdot) &= \textrm{the indicator function} \\
N\ &= \textrm{the number of individual projects that are being included in the analysis} \\
Pr_{p} &= \textrm{probability output from the algorithm for project}\ p \textrm{. Values are}\\
& \qquad \textrm{betweeen 0 and 1 where 1 is loss making)} \\
threshold &= \textrm{a chosen value between 0 and 1.}\ \Delta\ Overall\ Profit\ \textrm{is calculated for} \\
& \qquad \textrm{several}\ threshold\ \textrm{values which defines the profit curve} \\
Profit_p &= \textrm{profit for individual project}\ p
\end{aligned}
$$

If the threshold were zero, all jobs were rejected, the profit would be $0, and the ratio to historical profits would be 0%. If the threshold was 1.0, all jobs were accepted and the profit would be the same as historical profits, giving a ratio of 100%. The total profits were plotted for each threshold and joined to make a profit curve. The aim was to find the optimal threshold point where mostly loss-making projects were rejected, resulting in higher profits. 100 profit curves were built for each method to achieve a power of 0.8 and a 95% confidence interval could also be determined around each threshold point. The final expected increase in profit combined with the percentage of projects to be rejected presents a clear scenario that the case study business managers could strategically assess for implementation.

 4. Prediction Results and Discussion
====================================

Present results clearly and concisely
discussion should explore the signficance of the results of the work, not repeat them.

#### 4.1 Individual Models

The violin plot below summarises the distributions of the 100 AUC values produced by each method. The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05. Boosted Trees was chosen as the comparison method because it was expected to perform well.

```{r mice_violin, fig.pos = 'H', fig.height = 5.5}

mice.totAUC <- readRDS("mice_df_auc.rds")

#significant difference??
index = c(1,3,4)
pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
for(i in seq_along(index)){
        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
                format(digits = 2) %>% as.numeric
}

signif = pvals %>% filter(p_value < 0.05) %>% .$method
mice.totAUC = melt(mice.totAUC)
mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)

ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
        geom_violin(aes(fill = sigdiff, colour = sigdiff), 
                    alpha = 0.8) + 
        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
                     geom="text", colour = "gray20", size =3, 
                     hjust = -.2, angle = 40, vjust = -.2
                     ) +
        scale_fill_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly different from boosted trees") +
        scale_colour_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly different from boosted trees") +
        labs(x = "Method", y = "Area Under Curve (AUC)",
             title = "AUC Distribution from 100 Models of Each Method") +
        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
                         "log" = "Logistic Regression",
                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
        theme(legend.position = "bottom")
```

The AUC distrubtions of Logistic Regression and the Random Forest algorithms could not be statistically differentiated from Boosted Trees, while the Naive Bayes distribution was significantly lower. The three high-performing methods had AUC values between 0.755 and 0.764. This indicates predictions were well above random chance, where AUC = 0.5.

Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these. Therefore it was surprising that results from 100 Logistic Regression models were not statistically significantly lower than Boosted Trees. A possible explanation for this, according to the literature, is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel. Trees tend to overfit the patterns in a smaller training set. Logistic Regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003]. This may explain Logistic Regression's comparatively high performance on the case study's data set, which is small for the range and complexity of the described projects.

#### 4.2 Blended Models

Six blended models combined the predictions from the three high-performing individual models: Logistic Regression, Random Forests, and Boosted Trees to create 'blended' predictions. There were 3 simple blended models, and 3 complex models that interacted original covariates with the individual model predictions. Through variable importance analysis on the original covariates and the 3 new variables containing model predictions, only 6 original covariates were included as metafeatures. All six blended methods' AUC values were compared against the Logistic Regression AUC distribution, which was chosen because it was the simplest high performing individual model.

```{r blend_violin, fig.height = 5.5, fig.pos = 'h'}

auc.list <- readRDS("blend_auc2.rds")
means = lapply(auc.list, mean) %>% unlist
greater = means[means> means["orig.log"]] %>% names
auc.melt = melt(auc.list)
auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)

ggplot(auc.melt, aes(x= factor(L1), y= value)) +
        geom_violin(aes(fill = greater, colour = greater),
                    alpha = 0.8) +
        stat_summary(fun.y="mean", geom="point", colour = "gray30",
                     shape =4) +
        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
                     geom="text", colour = "gray20", size =3,
                     hjust = -.2, angle = 40, vjust = -.2
                     ) +
        theme(axis.text.x=element_text(angle=40,hjust=1)) +
        labs(x = "Method", y = "Area Under Curve (AUC)",
             title = "AUC Distribution from 100 Blended Models of each Method") +
        scale_x_discrete(labels = c("boost" = "BT Blend",
                         "log" = "FWLS",
                         "average" = "Simple Average Blend",
                         "orig.log" = "Orignal LR",
                         "rf" = "RF Blend",
                         "simp.boost" = "Simple BT Blend",
                         "simp.log" = "Simple LR Blend")) +
        scale_fill_manual(values = c('navajowhite','sienna1'),
                          name = "AUC Greater than Original Logistic Regression") +
        scale_colour_manual(values = c('navajowhite','sienna1'),
                          name = "AUC Greater than Original Logistic Regression") +
        theme(legend.position = "bottom")

```

None of the blended methods were significantly higher than the individual Logistic Regression model with data from 100 models. However, the simple averaged method and simple logistic regression could achieve a statistical power of 0.8 with 150 samples (models), whereas the other blended methods required thousands. 50 more models were run for the simple logistic regression and simple average blend to confirm their AUC distributions were significantly higher than the individual Logistic Regression. P-values from two sample t-test were 0.0019 and 0.0021 respectively.

Blended models improved the mean AUC from 0.759 to 0.77, which is an increase of `r ((0.77-0.759)/0.759) %>% "*"(100) %>% round(1)`%. This confirms the expectation that model blending would improve predictions by combining the strengths of different methods' theoretical foundations.

Again, the simplest methods worked best. In this case, averaging the results of the best three individual models or taking a Logistic Regression of the three models' outputs outperformed the complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated meta-feature interactions. Simiarly to the individual models, Simple methods may outperform complex methods if the data is not big enough for complex methods to learn the intricate patterns at which they excel. The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of their complex blending methods (Feature Weighted Linear Stacking) [@Sill2009]

The outcomes in this section show that statistical and machine learning methods can predict whether a consulting project will be profitable or not to some extent. The average, or weighted average (logistic regression) blend of individual Logistic Regression, Boosted Trees, and Random Forests models produced the highest AUC values. However, the business value of the model has not been proven. The following section explores whether the models would improve overall profits for the case study business.

5. Business Impact
==================

This chapter transports model results into the value framework to assess impact on the case study's overall profitability. The plot below illustrates the distribution of profit curves for the 100 models built using each method. Solid lines join the mean values at each threshold point and the grey ribbon represents the 95% confidence interval for the profit ratio.

```{r profit_curve, fig.height = 6, fig.pos = 'h'}

finance.plot = readRDS("plot_finance_100_4.rds" )
finance.plot$method <- factor(finance.plot$method,
                                    levels=c("average", "simp.log","simp.boost",
                                             "comp.boost","comp.rf", "FWLS","orig.log",
                                             "orig.rf","orig.boost"))

ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) +
        # geom_point(alpha = 0.005, shape = 16, size = 1) +
        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) +
        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
        stat_summary(fun.y = mean,
                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
        stat_summary(fun.y = mean, geom="line") +
        scale_y_continuous(breaks = seq(0,110, by = 20)) +
        scale_x_continuous(breaks = seq(0,1, by = .2)) +
        theme(legend.position = "none",
              panel.background = element_rect(fill = 'gray95')) +
        labs(x = "Probability Threshold",
             y = "Ratio of Model Profit to Historical Profit (%)",
             title = "Profit Curve for Blended and Individual Models") +
        facet_wrap(~method,
                   labeller = labeller( method = c("average" = "Simple Average Blend",
                                       "orig.boost"= "Orig. Boosted Trees",
                                       "orig.log" = "Orig. Logistic Regression",
                                       "orig.rf" = "Orig. Random Forest",
                                       "simp.log" = "Simple Log. Reg. Blend",
                                       "simp.boost" = "Simple Boosted Tree Blend",
                                       "comp.boost"= "Complex Boosted Tree Blend",
                                       "comp.rf"= "Complex Random Forest Blend",
                                       "FWLS" = "FWLS"
                                       )))

```

From the profit curves, the optimal probability thresholds are where the profit curves peak, and peaks above 100% indicate the model assisted profits outperformed historical profits. The two simple blended models clearly outperformed all other models with higher profit ratios as well as tighter confidence intervals. The simple Logistic Regression blend performed best with the highest mean profit ratio of 109% and a standard deviation of 1.28%. This means that if all jobs above the probability threshold 0.6 were rejected, the profits would increase on average by 9% in comparison to historical profits. Only 4.3% of projects were rejected in this case.

Profit curves from the simple blended models outperformed the complex blended models, which follows logically from their significantly higher AUC distributions. The shaded grey 95% confidence intervals show that none of the lower bounds for the original methods' or complex blending methods were above 100% on the y-axis. That is, it cannot be said with 95% confidence that these methods would produce a mean profit higher than historical profits in the given decision framework. It is interesting that despite the reasonable AUC values, the profit curves for the individual and complex blended models do not make a complelling case for adoption into business practice.

The simple blended Logistic Regression model on the other hand, a mean 9% improvement in profits, with a lower bound of 6.5%. In terms of implementing the model, this improvement may be high enough to trigger further development and analysis of a more sophisticated framework describing decision making scenarios.

6. Conclusions and Future Work
==============================

3/4 page

Complex ensemble tree methods did not outpereform simple linear methods, however, the simple blended predictions of all methods

The general aim of this work was twofold. First, statistical and machine learning techniques were tested to determine whether the profitability of projects for a case study consulting business could be predicted using their internal CRM data. This was rigorously completed by approaching the prediction of 'profitability' as a binary classification problem with classes profititable or loss-making. Naive Bayes, Logistic Regression, Random Forests, gradient Boosted Trees as well as simple and complex blending methods were applied to the problem. Secondly, this project tested whether predictions from the models could be proven to improve the overall profitability of the case study company and by how much.

Predicting whether a job would be profitable or not (binary classification) was possible to a level significantly better than random assignment (AUC = 0.76). This mean statistic was achieved by 100 different training/testing divisions of the data, built for each of the 3 individual methods. Simple averaging and simple Logistic Regression blending methods improved results further to a mean of AUC = 0.77.

Two models could also be shown to improve the overall profitability of the case study business. This was done by converting the binary predictions from the first analysis into profit curves that plotted model-influenced profit improvements over a range of probability threshold values. To determine model-influenced profit improvements, a managerial decision framework was used where projects scored below the threshold were accepted, while projects above the threshold were rejected. If a project was rejected, the profits and losses were forfeited, and the remaining accepted profits and losses were summed.

100 training/testing divisions of the data were user to produce 100 profit curves so that a 95% confidence interval could be framed around the curves. Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most. The 95% confidence interval for these improvements was between 6.5% and 11.5% using a probability threshold of 0.6 (approximately 4.3% of projects were rejected).

These results contribute significantly to the research in cost estimation in two ways: the applied methods, and the value assessment based on a managerial decision framework. Ensemble tree methods and blending had been applied minimally to cost estimation previously, even though ensemble trees provide insight into model structure while predicting at a similar level to Neural Networks. Next, previous studies compared accuracy of their predictive models, but stopped short of computing how the algorithm would affect decisions and what the measured benefits would be. This study presented a clear framework for how a business could improve profits by applying the algorithm.

Further work is required to measure user confidence in the output. Another topic identified for future research is to test how well managers estimate input variables at project conception. This is paraticularly relevant for time span and total invoiced amount which were discretised into 4 to 5 categories.

Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, from a consulting company's internal project data. This blended model predicts whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business by a mean of 9%.

Acknowledgments
===============

This research was supported by a scholarship/ ACEMS?

Figure Captions
===============

1. mice_violin: Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss. Each method was fed the same imputed full dataset.
2. blend_violin: Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss. 100 models were built for each method.
3. profit_curve: Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.

References {#references .unnumbered}
==========

#### Installation

If the document class *elsarticle* is not available on your computer,
you can download and install the system package *texlive-publishers*
(Linux) or install the LaTeX package *elsarticle* using the package
manager of your TeX installation, which is typically TeX Live or MikTeX.

The author names and affiliations could be formatted in two ways:

(1) Group the authors per affiliation.

(2) Use footnotes to indicate the affiliations.

Bullet points.

-   document style

-   baselineskip

-   front matter

-   keywords and MSC codes